Structured Versus Decorated Cospans

### 2020-11-19 cospans.tex
monoidal structure of E1 looks like strict

$$x_1+x_2 \mapseqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2)$$}

@@

\begin{thm}\label{thm:SC}
Given categories $\X$ and $\A$ with finite colimits and $L \maps \A \to \X$ a functor preserving
finite colimits, there is a symmetric monoidal double category $_L\lCsp(\X)$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a cospan in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\end{displaymath}
\item a 2-morphism is a map of cospans in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{displaymath}
\end{itemize}
Composition is formed by taking pushouts in $\X$ and identity horizontal 1-cells are $L(a)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}L(a)\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}L(a)$. 
Moreover, the vertical category is cocartesian monoidal $(\A,+,0_A)$ and the tensor of horizontal 1-cells is
\begin{displaymath}
\begin{tikzcd}[column sep={.4in,between origins}, row sep=.1in]
& x &&&& x' &&&& x+x' & \\
&&&\ot&&&& = &&&& \\
L(a)\ar[uur,"i"]&& L(b)\ar[uul,"o"'] && L(a')\ar[uur,"i'"] && L(b')\ar[uul,"o'"'] && L(a+a')\ar[uur,"i+i'"] && L(b+b')\ar[uul,"o+o'"'] 
\end{tikzcd}
\end{displaymath}
using that $L$ preserves coproducts, with monoidal unit $L(0_A)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$\sim$}}0_X\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$\sim$}}L(0_A)$. %!!! {\chris Shouldn't the horizontal monoidal unit be the identity on the vertical monoidal unit?}
\end{thm}

@@

The following theorem establishes the isomorphism between structured and decorated cospan double categories under certain conditions.  We denote by $\Rex$ the 2-category of finitely cocomplete categories, finitely cocontinuous functors and natural transformations.   (In \cite{BC} we took as objects finitely cocomplete categories with chosen finite colimits, just so we could turn these into symmetric monoidal categories without any further choices!!!  Should we do that here???) {\chris Can you tell if we can get away without doing that?}

@@

Since both double categories are symmetric monoidal under the running assumptions (\cref{DC,thm:SC}), we will now show that this double isomorphism is a symmetric monoidal one, namely $\lE=(\lE_0,\lE_1)$ is a symmetric strong monoidal double functor as per \cref{defn:monoidal_double_functor}. Indeed, $\lE_0=\mathrm{id}_\A$ is a symmetric monoidal functor trivially

>>>>

\begin{equation}\label{eq:tensordecoration}
x_1+x_2 \mapseqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2)
\end{equation}}

@@

\begin{thm}\label{thm:SC}
Given categories $\X$ and $\A$ with finite colimits and $L \maps \A \to \X$ a functor preserving
finite colimits, there is a symmetric monoidal double category $_L\lCsp(\X)$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a cospan in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\end{displaymath}
\item a 2-morphism is a map of cospans in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{displaymath}
\end{itemize}
Composition is formed by taking pushouts in $\X$ and identity horizontal 1-cells are $L(a)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}L(a)\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}L(a)$. 
Moreover, the vertical category is cocartesian monoidal $(\A,+,0_A)$ and the tensor of horizontal 1-cells is
\begin{equation}\label{eq:tensorstuctured}
\begin{tikzcd}[column sep={.4in,between origins}, row sep=.1in]
& x &&&& x' &&&& x+x' & \\
&&&\ot&&&& = &&&& \\
L(a)\ar[uur,"i"]&& L(b)\ar[uul,"o"'] && L(a')\ar[uur,"i'"] && L(b')\ar[uul,"o'"'] && L(a+a')\ar[uur,"i+i'"] && L(b+b')\ar[uul,"o+o'"'] 
\end{tikzcd}
\end{equation}
using that $L$ preserves coproducts, with monoidal unit $L(0_A)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$\sim$}}0_X\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$\sim$}}L(0_A)$. {\chris Shouldn't the horizontal monoidal unit be precisely the identity on the vertical monoidal unit?}
\end{thm}

@@

The following theorem establishes the isomorphism between structured and decorated cospan double categories under certain conditions.  We denote by $\Rex$ the 2-category of finitely cocomplete categories, finitely cocontinuous functors and natural transformations.   (In \cite{BC} we took as objects finitely cocomplete categories with chosen finite colimits, just so we could turn these into symmetric monoidal categories without any further choices!!!  Should we do that here???) {\chris Can you tell if we can get away without doing that? Perhaps after you read up to before the proof?}

@@

Since both double categories are symmetric monoidal under the running assumptions (\cref{DC,thm:SC}), we will now show that this double isomorphism is a symmetric monoidal one, namely $\lE=(\lE_0,\lE_1)$ is a symmetric strong monoidal double functor as per \cref{defn:monoidal_double_functor}. Indeed, $\lE_0=\mathrm{id}_\A$ is a symmetric monoidal functor trivially. For $\lE_1$, given two structured cospans $M=(a,\bot_a)\to(m,x)\leftarrow(b,\bot_b)$ and $M'=(a',\bot_{a'})\to(m',x')\leftarrow(b',\bot_{b'})$ in $_L\lCsp(\inta F)$, we compute the appropriate decorated cospans
\begin{align*}
\lE(M)\ot\lE(M')&\;\stackrel{\cref{eq:tensordecoration}}{=}\;a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\mathbf{1}\xrightarrow{x\times x'}F(m)\times F(m')\xrightarrow{\phi_{m,m'}}F(m+m') \\
\lE(M\ot M')&\;\stackrel{\cref{eq:tensorstuctured}}{=}\;\lE\Big((a+a',\bot_{a+a'})\to\underbrace{(m+m',\phi_{m,m'}(x,x'))}_{\cref{eq:explicitstructure2}}\leftarrow(b+b',\bot_{b+b'})\Big)%=a+a'\xrightarrow{i+i'}m+m'\xleftarrow{o+o'}b+b' \textrm{ with }\phi_{m,m'}(x,x')
\end{align*}
which are essentially the same. {\chris Well they look equal to me...thoughts?} Moreover, the monoidal unit for the category of arrows of $F\lCsp$ is $0_\A\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}0_\A\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}0_\A$ with the trivial decoration
$$\mathbf{1}\xrightarrow{\phi_0}F(0_A)\xrightarrow{F(\mathrm{id})}F(0_A)$$
which is essentially the same as $\lE((0_A,\phi_0)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}(0_A,\phi_0)\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}(0_A,\phi_0))$ again by \cref{eq:explicitstructure2}.
